Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-03-26 article-models-ttmodels.tex
new compatibility rules

>>>>

\subsubsection{Families of extended models}
Suppose we have a model $\ctxext{A}{P}$ in context $\Gamma$. In this section we
will figure out what the families of this model are. This is a bit tricky, so
we first treat the case where $P\jdeq\mfam{A}$. A context $\gamma$ in this
model is decomposed as a pair $\tmext{A}{\mfam{A}}{\pts{\gamma}}{\edg{\gamma}}$
which consists of a context $\pts{\gamma}$ of the model $A$ and a context
$\edg{\gamma}$ of the model $\mfam{A}$ in the fiber above $\pts{\Gamma}$. In
other words, $\edg{\gamma}$ is a family over $\pts{\gamma}$. A family $\beta$
over $\gamma$ is then a pair $\tmext{}{}{\pts{\beta}}{\edg{\beta}}$ consisting
of a family $\pts{\beta}$ over $\pts{\gamma}$ and a family $\edg{\beta}$ over
the extended term $\tmext{}{}{\pts{\beta}}{\edg{\gamma}}$. 
It is this extension which appears in the type of $\edg{\beta}$ which requires
a workaround for general $P$.

We consider the following diagram helpful:
\begin{equation*}
\begin{tikzcd}
\jcomp{}{\bar{\typefont{pr}}}{{}{\tfext{A}}{\mfam{P}}}
  \rar{}
  \ar[fib]{d}
  &
  \jcomp{}{\tfext{A}}{\mfam{P}}
    \rar{}
    \ar[fib]{d}{}
  &
  \mfam{P}
    \ar[fib]{d}{}
  \\
\jcomp{}{\bar{\typefont{pr}}}{{}{\tfext{A}}{P}}
  \rar{}
  \ar[fib]{d}{}
  &
  \jcomp{}{\tfext{A}}{P}
  \rar{}
  \ar[fib]{d}{}
  &
  P
  \ar[fib]{d}{}\\
\ctxext{{A}{P}}{\mfam{A}}
  \ar{r}[swap]{\bar{\typefont{pr}}}
  &
  \ctxext{A}{\mfam{A}}
  \ar{r}[swap]{\tfext{A}}
  &
  A
\end{tikzcd}
\end{equation*}
Here $\bar{\typefont{pr}}\jdeq\tmext{}{}{\cprojfstf{A}{P}}{\idtm{\mfam{A}}}$,
the unmarked horizontal arrows indicate identity terms and the vertical arrows
merely indicate that what's above is a family depending on what's below.

\subsubsection{Families of substituted modes}
The following inference rule explains that the fiber at a term $f$ of $P$
of the model of families of a family of models $Q$ over $\ctxext{{\Gamma}{A}}{P}$
is judgmentally the same as the model of families of the fiber at the term $f$
of the family $Q$. In more practical terms, suppose we have a context $\gamma$
in $Q$ which happens to be in the fiber at $f$ of $Q$ over $P$, then a family
over $\gamma$ according to $Q$ is the same thing as a family over $\Gamma$
according to $\subst{f}{Q}$. The following inference rule captures this idea
precisely:

\begin{equation}
\inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jfameq
      {{\Gamma}{A}}
      {\subst{f}{\mfam{Q}}}
      {\mfam{\subst{f}{Q}}}
    }
\end{equation}

\subsubsection{Families of weakened models}
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \jfameq
      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\ctxwk{A}{\mfam{B}}}
      {\mfam{\ctxwk{A}{B}}}
    }
\end{equation*}

\subsubsection{The action on families of the identity term}
Let $A$ be a model in context $\Gamma$. Then we have the identity term
$\jhom{\Gamma}{A}{A}{\idtm{A}}$, which also has an action on families with the
following specification:
\begin{equation*}
\jhom{{\Gamma}{A}}{\mfam{A}}{\subst{\idtm{A}}{\mfam{\ctxwk{A}{A}}}}{\mfam{\idtm{A}}}.
\end{equation*}
Note, however that we have the judgmental equalities
\begin{equation*}
\subst{\idtm{A}}{\mfam{\ctxwk{A}{A}}}\jdeq\subst{\idtm{A}}{\ctxwk{A}{\mfam{A}}}
\jdeq\mfam{A},
\end{equation*}
so we may require that the action on families of the identity
term $\idtm{A}$ is the identity term $\idtm{\mfam{A}}$. We get the following
inference rule:
\begin{equation*}
\inference
  { \jfam{\Gamma}{A}
    }
  { \jhomeq
      {{\Gamma}{A}}
      {\mfam{A}}
      {\mfam{A}}
      {\mfam{\idtm{A}}}
      {\idtm{\mfam{A}}}
    }
\end{equation*}
